---
title: "HW2: OLS and Linear Algebra"
author: "Kaori hirano"
date: "6 14 2023"
format: pdf
---

# Packages

```{r load-packages}
# load packages here
library(plyr)
library(tidyverse)
install.packages('lmSupport')
library(lmSupport)
```

# Data

```{r import-data}
# import data here
dcbikeshare <- read.csv('data/bikeshare-day.csv')
```


# Data Wrangling

## Q1 
Recode the season variable to be a factor with meaningful level names as outlined in the
codebook (at the link above), with “spring” as the baseline level.

```{r q1}
# code here
d <- dcbikeshare
as.factor(d$season)
(d$season <- mapvalues(d$season, from = c('1','2','3', '4'), to = c("Winter", "Spring", "Summer", "Fall"))) # recodes, not ideal method but lmSupport wasn't loading...

d$season <- fct_relevel(d$season, c('Spring', 'Summer', 'Fall', 'Winter')) #relevels
levels(d$season)
```

## Q2 
Recode the binary variables holiday and workingday to be factors with levels “no” (0) and
“yes” (1), with “no” as the baseline level.
```{r q2}
# recodes 0 to no and 1 to yes
d$hbin <- mapvalues(d$holiday, from = c('0', '1'), to = c('no', 'yes'))
# cbind(head(d$hbin), head(d$holiday)) # checks accuracy 
d$workbin <- mapvalues(d$workingday, from = c('0', '1'), to = c('no', 'yes')) 
# cbind(d$workbin, d$workinday) # checks accuracy 
```

## Q3
Recode the yr variable to be a factor with levels 2011 and 2012, with 2011 as the baseline
level.
```{r q3}
d$years <- mapvalues(d$yr, from = c('0', '1'), to = c('2011', '2012'))
# cbind(d$years, d$yr) # checks accuracy 
```

## Q4
Recode the weathersit variable as 1 - “clear”, 2 - “mist”, 3 - “light precipitation,” and 4 -
“heavy precipitation”, with “clear” as the baseline
```{r q1}
# code here
as.factor(d$weathersit)

# d <- d %>%
#  mutate(weathersit = case_when(
#    weathersit == "1"   ~ "clear",
#    weathersit == "2"   ~ "mist",
#    weathersit == '3'   ~ "light precipitation",
#    weathersit == "4"   ~ "heavy precipitation",
#    TRUE                ~ weathersit
#  ))
```

## Q5
Calculate raw temperature, feeling temperature, humidity, and windspeed as their values given
in the dataset multiplied by the maximum raw values stated in the codebook for each variable.
Instead of writing over the existing variables, create new ones with concise but
informative names.
```{r q1}
# multiplies normalized values by their stated maximum values to get raw values
d$raw_temp <- d$temp * 41
d$raw_ftemp <- d$atemp * 50
d$raw_hum <- d$hum * 100
d$raw_ws <- d$windspeed * 67
```

# Data Visualization 

## Q6

```{r data-viz}
# recreases image with raw feeling temp
ggplot(d, aes(x = dteday, y = cnt, color = raw_ftemp)) + 
  geom_point() +
  labs(title = "Bike rentals in DC, 2011 and 2012",
       subtitle = "Warmer temperatures associated with more bike rentals",
       x = "Date", y = "Bike rentals",
       color = "Temperature (C)",
       caption = "Source: Palmer Station LTER / palmerpenguins package") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   scale_x_continuous(breaks = c(, 20), labels = c("2011-01", "2011-07", '2021-01', '2021-07', '2013-01'))
# find a way to get these specific dates
```

*Write narrative response here.*

# Modeling

## Q7
```{r q7}
lm1 <- lm(cnt ~ raw_temp, d) # num rentals by raw temperature
summary(lm1) 

temp <- (63 - 32) * (5/9) # converts to C
(pred <- 1214.642 + 161.969*temp) # puts into equation
confint(lm1) # gets CI

new <- data.frame(raw_temp = (temp)) # makes dataframe for predict

predict(lm1, newdata = new, interval = 'confidence')
predict(lm1, newdata = new, interval = 'predict')
```
a) There is a relationship between the total daily bike rentals and the daily temperature (b = 161.969, F(1,729) = 473.5, p < .0001)

b) The relationship seems to be moderate as seen by the adjusted R squared of .3929. 

c) The relationship is positive as indicated by the positive R squared. 

d) The predicted number of bike rentals with a temp of 63 degrees F is 4004. The associated 95% confidence interval is (3885.57, 4122.63) and prediction interval is (1038.464, 6969.737).

## Q8
```{r q8}
lm2 <- lm(cnt ~ raw_ftemp, d) # num rentals by raw temperature
summary(lm2) 
summary(lm1)
```
Feeling temperature is a better predictor of bike rentals, although not by large amount. Both have a statistically significant result (pval < .001), but there is a lower standard error for feeling temperature and more importantly a higher correlation coefficient and F-statistic. 

## Q9
```{r q9}
# fit full model
names(d)
lm_full <- lm(cnt ~ season + years + holiday + workingday + weathersit + raw_temp + raw_ftemp + raw_hum + raw_ws + (raw_ftemp * holiday), d)
summary(lm_full)
```
Yes, there is a relationship between the predictors and the response, as seen by the adjusted R-squared value of .8142 indicating a high level of correlation. 

b) the predictors that appear to have a statistically significant relationship to the amount of bike rentals are the season, the year, holiday status, weather type, raw temperature, raw humidity, and raw windspeed. 

c) the coefficients for the season suggest that there is a negative relationship for summer and winter with bike rentals and a positive relationship with fall when compared to a baseline level of spring's effect on rentals. 

# Using Linear Algebra to Do Regression

## Q10
```{r q10}
X <- model.matrix(lm_full)
```

## Q11 refer to q9
```{r q11}
y <- model.frame(lm_full)
```

## Q12
```{r q12}
(vec <- (solve(t(X)*X)*t(X)*y))
```
