---
title: "HW2: OLS and Linear Algebra"
author: "Kaori hirano"
date: "6 14 2023"
format: pdf
---

# Packages

```{r load-packages}
# load packages here
library(plyr)
library(tidyverse)
install.packages('lmSupport')
library(lmSupport)
```

# Data

```{r import-data}
# import data here
dcbikeshare <- read.csv('data/bikeshare-day.csv')
```


# Data Wrangling

## Q1 
Recode the season variable to be a factor with meaningful level names as outlined in the
codebook (at the link above), with “spring” as the baseline level.

```{r q1}
# code here
d <- dcbikeshare
as.factor(d$season)
(d$season <- mapvalues(d$season, from = c('1','2','3', '4'), to = c("Winter", "Spring", "Summer", "Fall"))) # recodes, not ideal method but lmSupport wasn't loading...

d$season <- fct_relevel(d$season, c('Spring', 'Summer', 'Fall', 'Winter')) #relevels
levels(d$season)
```

## Q2 
Recode the binary variables holiday and workingday to be factors with levels “no” (0) and
“yes” (1), with “no” as the baseline level.
```{r q2}
# recodes 0 to no and 1 to yes
d$hbin <- mapvalues(d$holiday, from = c('0', '1'), to = c('no', 'yes'))
# cbind(head(d$hbin), head(d$holiday)) # checks accuracy 
d$workbin <- mapvalues(d$workingday, from = c('0', '1'), to = c('no', 'yes')) 
# cbind(d$workbin, d$workinday) # checks accuracy 
```

## Q3
Recode the yr variable to be a factor with levels 2011 and 2012, with 2011 as the baseline
level.
```{r q1}
# code here
```

## Q4
Recode the weathersit variable as 1 - “clear”, 2 - “mist”, 3 - “light precipitation,” and 4 -
“heavy precipitation”, with “clear” as the baseline
```{r q1}
# code here
```

## Q5
Calculate raw temperature, feeling temperature, humidity, and windspeed as their values given
in the dataset multiplied by the maximum raw values stated in the codebook for each variable.
Instead of writing over the existing variables, create new ones with concise but
informative names.
```{r q1}
# code here
```

# Data Visualization 

## Q6

```{r data-viz}
# code to recreate plot here
```

*Write narrative response here.*

# Modeling

## Q7
```{r q1}
# code here
```

## Q8
```{r q1}
# code here
```

## Q9
```{r q1}
# code here
```

# Using Linear Algebra to Do Regression

## Q10
```{r q1}
# code here
```

## Q11
```{r q1}
# code here
```

## Q12
```{r q1}
# code here
```
